Ellipsoids (v1.0): 3D magnetic modelling of ellipsoidal bodies

### 2017-09-25 manuscript/manuscript.tex
final proofreading version

\title{Ellipsoids (v1.0): 3D Magnetic modelling of ellipsoidal bodies}

@@

\Author[1]{Diego}{Takahashi Tomazella}
\Author[1]{Vanderlei}{C. Oliveira Jr.}

@@

\runningauthor{Tomazella and Oliveira Jr.}

@@

A considerable amount of literature has been published on the magnetic
modelling of uniformly magnetized ellipsoids since the second half of
the nineteenth century. Ellipsoids have flexibility to represent a wide
range of geometrical forms, are the only known bodies which can be
uniformly magnetized in the presence of a uniform inducing field and
are the only finite bodies for which the self-demagnetization can be treated
analytically. This property makes ellipsoids particularly useful for
modelling compact orebodies having high susceptibility. In this case,
neglecting the self-demagnetization may strongly mislead the interpretation
of these bodies by using magnetic methods. A number of previous studies
consider that the self-demagnetization can be neglected for the case in
which the geological body has an isotropic susceptibility lower than or
equal to 0.1 SI. This limiting value, however, seems to be determined
empirically and there has been no discussion about how this value was
determined. Besides, the geoscientific community lacks an easy-to-use
tool to simulate the magnetic field produced by uniformly magnetized
ellipsoids. Here, we present an integrated review of the magnetic
modelling of arbitrarily oriented triaxial, prolate and oblate
ellipsoids. Our review includes ellipsoids with both induced and
remanent magnetization, as well as with isotropic or anisotropic
susceptibility. We also discuss the ambiguity between confocal ellipsoids
with the same magnetic moment and propose a
way of determining the isotropic susceptibility above which the
self-demagnetization must be taken into consideration. Tests with
synthetic data validate our approach. Finally, we provide a set
of routines to model the magnetic field produced
by ellipsoids. The routines are written in Python language as
part of the Fatiando a Terra, which is an open-source library
for modelling and inversion in geophysics.

@@

Based on the mathematical theory of the magnetic
induction developed by \citet{poisson1824},
\citet{maxwell1873} affirmed that, if $U$ is the
gravitational potential produced by any body
with uniform density $\rho$ and arbitrary shape at
a point $(x, y, z)$, then $-\frac{\partial U}{\partial x}$
is the magnetic scalar potential produced
at the same point by the same body
if it has a uniform magnetization oriented along $x$
with intensity $\rho$.
\citet{maxwell1873} generalized this idea as a way
of determining the magnetic scalar potential produced
by any uniformly magnetized body in a given direction.
By presuming that this uniform magnetization is due to
induction and that it is proportional to the resulting
magnetic field (intensity) inside the body, he postulated
that the resulting field must also be uniform and
parallel to the magnetization. This uniformity is due to
the fact that the resulting field is defined as the
negative gradient of the magnetic scalar potential.
As a consequence of this uniformity,
the gravitational potential $U$ at points within the
body must be a quadratic
function of the spatial coordinates.
Apparently, \citet{maxwell1873} was the first one to
postulate that ellipsoids are the only finite bodies
having a gravitational potential which satisfies this
property and hence can be uniformly magnetized
in the presence of a uniform inducing magnetic field.
This property can be extended to other bodies
defined as limiting cases of an ellipsoid
(e.g., spheres, elliptic cylinders), however all the
remaining non-ellipsoidal bodies cannot be
uniformly magnetized in the presence of a uniform
inducing field.

Another particularity of ellipsoids is that they are
the only bodies which enable an analytical computation
of its self-demagnetization.
The self-demagnetization contributes to decrease the
magnitude of the magnetization along the shortest
axes of a body.
It is a function of the body shape and gives rise
to shape anisotropy \citep{uyeda1963, thompson1986,
dunlop1997, clark1999, tauxe2003}.
It is well-established in the literature that
the self-demagnetization can be neglected if the
body has a susceptibility lower than 0.1 \unit{SI}
\citep{emerson1985, clark1986, eskola1980, guo1998,
guo2001, purss2005, hillan2013, austin2014, clark2014}.
On the other hand, neglecting the self-demagnetization in
geological bodies with high susceptibilities (> 0.1 \unit{SI})
may strongly mislead the interpretation obtained from
magnetic methods.
This limiting value, however, seems to be determined empirically
and, so far, there has been little discussion about
how it was determined.

\citet{farrar1979} demonstrated the importance of the
ellipsoidal model in taking into account the
self-demagnetization and determining reliable
drilling directions on the Tennant Creek field,
Australia.
Posteriorly, \citet{hoschke1991} also showed how the
ellipsoidal model proved to be highly successful in
locating and defining ironstone bodies in the
Tennant Creek field.
\citet{clark2000} provides a good discussion about the
influence of the self-demagnetization in magnetic
interpretation of the Osborne copper-gold deposit,
Australia. This deposit is hosted by ironstone
bodies that have very high susceptibility.
According to \citet{clark2000}, neglecting the effects
of self-demagnetisation led errors of $\approx 55^{\circ}$
in the interpreted dip.
Recently, \citet{austin2014} used magnetic modelling
and rock property measurements to show that,
contrary to previous interpretations, the magnetization
of the Candelaria iron oxide copper-gold deposit,
Chile, is not dominated by the induced component.
Rather, the deposit has a relatively weak remanent
magnetization and is strongly affected by
self-demagnetization.
These examples show the importance of the self-demagnetization
and the ellipsoidal model in
producing trustworthy geological models of
high-susceptibility orebodies, which may save significant
cost associated with drilling.

A vast literature about the magnetic
modelling of ellipsoidal bodies was developed in which
are to be found the names of many researchers.
Nevertheless, interest in this subject has not yet died
out, as is evidenced by a list of modern papers in this
field.
Besides, the geoscientific community lacks a free easy-to-use
tool to simulate the magnetic field produced by uniformly
magnetized ellipsoids.
Such a tool could prove to be useful either for teaching
and researching geophysics.

In this work, we present a review of the vast literature
about the magnetic modelling of ellipsoidal bodies and a
theoretical discussion about the determination of the
isotropic susceptibility value above which
the self-demagnetization must be taken into consideration.
We propose an alternative way of determining this value
based on the body shape and the maximum relative error
allowed in the resultant magnetization.
This alternative approach is validated
by the results obtained with numerical simulations.
We also provide a set of routines to
model the magnetic field produced by ellipsoids.
The routines are written in Python language as part of
the Fatiando a Terra \citep{uieda2013}, which is an
open-source library for modelling and inversion in geophysics.
We attempt to use the best practices of continuous
integration, documentation, unit-testing, and version-control
for the purpose of providing a reliable and easy-to-use
code.

@@

Let $(x, y, z)$ be a point referred to a Cartesian coordinate system with axes
$x$, $y$ and $z$ pointing to, respectively, North, East and down.
For convenience, we denominate this coordinate system as
\textit{main coordinate system} (Fig. \ref{fig:orientation-angles}a).
Let us consider an ellipsoidal body with centre at the point
$(x_{c}, y_{c}, z_{c})$, orientation defined by the angles
\textit{strike} $\varepsilon$, \textit{dip} $\zeta$, and \textit{rake} $\eta$
(Fig. \ref{fig:orientation-angles}a), and semi-axes defined by
positive constants $a$, $b$, $c$ (Fig. \ref{fig:orientation-angles}b).
The orientation angles \textit{strike}, \textit{dip}, and \textit{rake}
are commonly used to define the orientation of lines in structural geology
\citep{pollard2005, allmendinger2012}.
The points $(x, y, z)$ located on the surface of this ellipsoidal body
satisfy the following equation:
\begin{equation}
(\mathbf{r} - \mathbf{r}_c)^T \mathbf{A} (\mathbf{r} - \mathbf{r}_c) = 1 \: ,
\label{eq:ellipsoid_surface}
\end{equation}
where $\mathbf{r} = [\begin{array}{ccc} x & y & z \end{array} ]^{\top}$,
$\mathbf{r}_{c} = [\begin{array}{ccc} x_{c} & y_{c} & z_{c} \end{array} ]^{\top}$,
$\mathbf{A}$ is a positive definite matrix given by
\begin{equation}
\mathbf{A} = \mathbf{V}
\left[ \begin{array}{ccc}
a^{-2} & 0 & 0 \\
0 & b^{-2} & 0 \\
0 & 0 & c^{-2}
\end{array} \right] \mathbf{V}^{\top} \: ,
\label{eq:A}
\end{equation}
and $\mathbf{V}$ is an orthogonal matrix whose first, second and third columns
are defined by unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$
(Fig. \ref{fig:orientation-angles}b), respectively.
The matrix $\mathbf{V}$ can be defined in terms of three rotation matrices:
\begin{equation}
\mathbf{R}_{1}(\theta) = \left[ \begin{array}{ccc}
1 & 0 & 0 \\
0 & \cos \theta & \sin \theta \\
0 & -\sin \theta & \cos \theta
\end{array} \right] \: ,
\label{eq:R1}
\end{equation}
\begin{equation}
\mathbf{R}_{2}(\theta) = \left[ \begin{array}{ccc}
\cos \theta & 0 & -\sin \theta \\
0 & 1 & 0 \\
\sin \theta & 0 & \cos \theta
\end{array} \right]
\label{eq:R2}
\end{equation}
and
\begin{equation}
\mathbf{R}_{3}(\theta) = \left[ \begin{array}{ccc}
\cos \theta & \sin \theta & 0 \\
-\sin \theta & \cos \theta & 0 \\
0 & 0 & 1
\end{array} \right] \: .
\label{eq:R3}
\end{equation}
For triaxial ellipsoids (i.e., $a > b > c$) and prolate ellipsoids
(i.e., $a > b = c$), we define the orthogonal matrix $\mathbf{V}$
as follows:
\begin{equation}
\mathbf{V} =
\mathbf{R}_{1} \left( \frac{\pi}{2} \right) \;
\mathbf{R}_{2} \left( \varepsilon \right) \;
\mathbf{R}_{1} \left( \frac{\pi}{2} - \zeta \right) \;
\mathbf{R}_{3} \left( \eta \right) \: .
\label{eq:V_triaxial_prolate}
\end{equation}
For oblate ellipsoids (i.e., $a < b = c$), we define $\mathbf{V}$ as follows:
\begin{equation}
\mathbf{V} =
\mathbf{R}_{3} \left( -\frac{\pi}{2} \right) \;
\mathbf{R}_{1} \left( \pi \right) \,
\mathbf{R}_{3} \left( \varepsilon \right) \,
\mathbf{R}_{2} \left( \frac{\pi}{2} - \zeta \right) \,
\mathbf{R}_{1} \left( \eta \right) \: .
\label{eq:V_oblate}
\end{equation}
The orthogonal matrices $\mathbf{V}$ used here for triaxial, prolate and oblate
ellipsoids (Eqs. \ref{eq:V_triaxial_prolate} and \ref{eq:V_oblate}) are
different from that used by \citet{emerson1985} and \citet{clark1986}.

The magnetic modelling of an ellipsoidal body is commonly performed
in a particular Cartesian coordinate system that is aligned
with the body semi-axes
and has the origin coincident with the body centre
(Fig. \ref{fig:orientation-angles}b).
For convenience, we denominate this particular coordinate
system as \textit{local coordinate system}.
The relationship between the Cartesian coordinates
$(\tilde{x}, \tilde{y}, \tilde{z})$ of a point in
a local coordinate system and the Cartesian
coordinates $(x, y, z)$ of the same point in the main
system is given by:
\begin{equation}
\tilde{\mathbf{r}} = \mathbf{V}^{\top} \left( \mathbf{r} - \mathbf{r}_{c} \right) \: ,
\label{eq:coord_transformation}
\end{equation}
where
$\tilde{\mathbf{r}} = [\begin{array}{ccc} \tilde{x} &
                                          \tilde{y} &
                                          \tilde{z} \end{array} ]^{\top}$,
$\mathbf{r}$ and $\mathbf{r}_{c}$
are defined in Eq. \ref{eq:ellipsoid_surface} and the matrix $\mathbf{V}$
(Eq. \ref{eq:A}) is defined according to the ellipsoid type.
Subsequently, quantities referred to the local coordinate system
(Fig. \ref{fig:orientation-angles}b) are represented with
the simbol "$\sim$".

@@

Consider a magnetized ellipsoid immersed in
a uniform inducing magnetic field $\mathbf{H}_{0}$ (in $\unit{Am^{-1}}$)
given by
\begin{equation}
\mathbf{H}_{0} = \| \mathbf{H}_{0} \| \left[
\begin{array}{c}
\cos I \: \cos D \\
\cos I \: \sin D \\
\sin I
\end{array}
\right] \: ,
\label{eq:H0}
\end{equation}
where $\| \cdot \|$ denotes the Euclidean norm (or 2-norm) and
$D$ and $I$ are respectively, the declination and inclination of the
local-geomagnetic field in the main coordinate system
(Fig. \ref{fig:orientation-angles}a).
This field represents the main component of the
Earth's magnetic field, which is usually assumed to be generated
by the Earth's liquid core.
In the absence of conduction currents,
the total magnetic field $\mathbf{H}(\mathbf{r})$
at the position $\mathbf{r}$ (Eq. \ref{eq:ellipsoid_surface})
of a point referred to the main
coordinate system is defined as follows \citep{sharma1966,
eskola1980, reitz1992, sttraton2007}:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0} - \nabla V(\mathbf{r}) \: ,
\label{eq:H}
\end{equation}
where the second term is the negative gradient of
the magnetic scalar potential $V(\mathbf{r})$ given by:
\begin{equation}
V(\mathbf{r}) = -\frac{1}{4\pi} \iiint_{\vartheta}
\mathbf{M}(\mathbf{r}^{\prime})^{\top}
\nabla \left(
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\right) \, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:V-potential}
\end{equation}
In this equation, $\mathbf{r}^{\prime} = [\begin{array}{ccc}
x^{\prime} & y^{\prime} & z^{\prime} \end{array} ]^{\top}$
is the position vector of a point located within the volume $\vartheta$,
the integral is conducted over the variables
$x^{\prime}$, $y^{\prime}$ and, $z^{\prime}$ and
$\mathbf{M}(\mathbf{r}^{\prime})$ is the magnetization vector
(in \unit{Am^{-1}}).
Eq. \ref{eq:V-potential} is valid anywhere,
independently if the position vector $\mathbf{r}$ represents
a point located inside or outside the magnetized body
\citep{dubois1896, reitz1992, sttraton2007}.

Based on \citeauthor{maxwell1873}'s postulate, let us
assume that the body has a uniform magnetization given by
\begin{equation}
\mathbf{M} = \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:M-KH}
\end{equation}
where $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field
at any point within the body and $\mathbf{K}$ is a
constant and symmetrical 2nd-order tensor representing the
magnetic susceptibility of the body.
This is a good approximation for bodies at room temperature,
subjected to an inducing field $\mathbf{H}_{0}$ with
strength $\leq 10^{-3} \mu_{0}^{-1}$ \unit{A \, m^{-1}} \citep{rochette1992},
where $\mu_{0}$ represents the magnetic constant (in \unit{H \, m^{-1}}).
In this case, the susceptibility tensor $\mathbf{K}$ is commonly
represented, in the main coordinate system (Fig. \ref{fig:orientation-angles}a),
as follows:
\begin{equation}
\mathbf{K} = \mathbf{U}
\left[ \begin{array}{ccc}
k_{1} & 0 & 0 \\
0 & k_{2} & 0 \\
0 & 0 & k_{3}
\end{array} \right] \mathbf{U}^{\top} \: ,
\label{eq:K}
\end{equation}
where $k_{1} > k_{2} > k_{3}$ are the
\textit{principal susceptibilities} and $\mathbf{U}$ is
an orthogonal matrix whose columns $\mathbf{u}_{i}$,
$i = 1, 2, 3$, are unit vectors called
\textit{principal directions}.
Similarly to the matrix $\mathbf{V}$ (Eqs. \ref{eq:ellipsoid_surface},
\ref{eq:V_triaxial_prolate} and \ref{eq:V_oblate}), we define
the matrix $\mathbf{U}$ as a function
of given orientation angles $\varepsilon$, $\zeta$,
and $\eta$ depending on the ellipsoid type. For triaxial
and prolate ellipsoids, we define $\mathbf{U}$ by
using Eq. \ref{eq:V_triaxial_prolate}, whereas for oblate ellipsoids
we use Eq. \ref{eq:V_oblate}.
Notice that the orientation angles $\varepsilon$, $\zeta$,
and $\eta$ defining the orthogonal matrix $\mathbf{U}$
may be different from that angles $\varepsilon$, $\zeta$,
and $\eta$ defining the ellipsoid orientation
(Fig. \ref{fig:orientation-angles}).

If the principal susceptibilities are different from
each other, we say that the body has an
anisotropy of magnetic susceptibility (AMS).
The AMS is generally associated to the preferred orientation
of the grains of magnetic minerals forming the rock
\citep{fuller1963, uyeda1963, janak1972, hrouda1982,
thompson1986, macdonald1987, rochette1992, dunlop1997,
tauxe2003}.
For the particular case in which the principal directions
coincide with the ellipsoid axes, the matrix $\mathbf{U}$ is
equal to the matrix $\mathbf{V}$ (Eq. \ref{eq:A}).
Another important particular case is that in which the
susceptibility is isotropic and, consequently, the principal
susceptibilities $k_{1}$, $k_{2}$, and $k_{3}$ (Eq. \ref{eq:K})
are equal to a constant $\chi$. In this case, the susceptibility
tensor $\mathbf{K}$ (Eq. \ref{eq:K}) assumes the particular form
\begin{equation}
\mathbf{K} = \chi \, \mathbf{I} \: ,
\label{eq:K-isotropic}
\end{equation}
where $\mathbf{I}$ represents the identity matrix.

>>>>

\title{Ellipsoids (v1.0): 3D magnetic modelling of ellipsoidal bodies}

@@

\Author[1]{Diego}{Takahashi}
\Author[1]{Vanderlei}{C. Oliveira Jr.}

@@

\runningauthor{D. Takahashi and V. C. Oliveira Jr.}

@@

A considerable amount of literature has been published on the magnetic
modelling of uniformly magnetized ellipsoids since the second half of
the nineteenth century. Ellipsoids have flexibility to represent a wide
range of geometrical forms, are the only known bodies which can be
uniformly magnetized in the presence of a uniform inducing field and
are the only finite bodies for which the self-demagnetization can be treated
analytically. This property makes ellipsoids particularly useful for
modelling compact orebodies having high susceptibility. In this case,
neglecting the self-demagnetization may strongly mislead the interpretation
of these bodies by using magnetic methods. A number of previous studies
consider that the self-demagnetization can be neglected for the case in
which the geological body has an isotropic susceptibility lower than or
equal to 0.1 SI. This limiting value, however, seems to be determined
empirically and there has been no discussion about how this value was
determined. In addition, the geoscientific community lacks an easy-to-use
tool to simulate the magnetic field produced by uniformly magnetized
ellipsoids. Here, we present an integrated review of the magnetic
modelling of arbitrarily oriented triaxial, prolate and oblate
ellipsoids. Our review includes ellipsoids with both induced and
remanent magnetization, as well as with isotropic or anisotropic
susceptibility. We also discuss the ambiguity between confocal ellipsoids
with the same magnetic moment and propose a
way of determining the isotropic susceptibility above which the
self-demagnetization must be taken into consideration. Tests with
synthetic data validate our approach. Finally, we provide a set
of routines to model the magnetic field produced
by ellipsoids. The routines are written in Python language as
part of the Fatiando a Terra, which is an open-source library
for modelling and inversion in geophysics.

@@

Based on the mathematical theory of the magnetic
induction developed by \citet{poisson1824},
\citet{maxwell1873} affirmed that, if $U$ is the
gravitational potential produced by any body
with uniform density $\rho$ and arbitrary shape at
a point $(x, y, z)$, then $-\frac{\partial U}{\partial x}$
is the magnetic scalar potential produced
at the same point by the same body
if it has a uniform magnetization oriented along $x$
with intensity $\rho$.
\citet{maxwell1873} generalized this idea as a way
of determining the magnetic scalar potential produced
by any uniformly magnetized body in a given direction.
By presuming that this uniform magnetization is due to
induction and that it is proportional to the resulting
magnetic field (intensity) inside the body, he postulated
that the resulting field must also be uniform and
parallel to the magnetization. This uniformity is due to
the fact that the resulting field is defined as the
negative gradient of the magnetic scalar potential.
As a consequence of this uniformity,
the gravitational potential $U$ at points within the
body must be a quadratic
function of the spatial coordinates.
Apparently, \citet{maxwell1873} was the first one to
postulate that ellipsoids are the only finite bodies
having a gravitational potential which satisfies this
property and hence can be uniformly magnetized
in the presence of a uniform inducing magnetic field.
This property can be extended to other bodies
defined as limiting cases of an ellipsoid
(e.g. spheres, elliptic cylinders). However all the
remaining non-ellipsoidal bodies cannot be
uniformly magnetized in the presence of a uniform
inducing field.

Another particularity of ellipsoids is that they are
the only bodies which enable an analytical computation
of their self-demagnetization.
The self-demagnetization contributes to a decrease in the
magnitude of the magnetization along the shortest
axes of a body.
This a function of the body shape and gives rise
to shape anisotropy \citep{uyeda1963, thompson1986,
dunlop1997, clark1999, tauxe2003}.
It is well established in the literature that
the self-demagnetization can be neglected if the
body has a susceptibility lower than 0.1 \unit{SI}
\citep{emerson1985, clark1986, eskola1980, guo1998,
guo2001, purss2005, hillan2013, austin2014, clark2014}.
On the other hand, neglecting the self-demagnetization in
geological bodies with high susceptibilities (> 0.1 \unit{SI})
may strongly mislead the interpretation obtained from
magnetic methods.
This limiting value, however, seems to be determined empirically
and, so far, there has been little discussion about
how it was determined.

\citet{farrar1979} demonstrated the importance of the
ellipsoidal model in taking into account the
self-demagnetization and determining reliable
drilling directions on the Tennant Creek field,
Australia.
Later, \citet{hoschke1991} also showed how the
ellipsoidal model proved to be highly successful in
locating and defining ironstone bodies in the
Tennant Creek field.
\citet{clark2000} provides a good discussion about the
influence of the self-demagnetization in magnetic
interpretation of the Osborne copper-gold deposit,
Australia. This deposit is hosted by ironstone
bodies that have very high susceptibility.
According to \citet{clark2000}, neglecting the effects
of self-demagnetisation led to errors of $\approx 55^{\circ}$
in the interpreted dip.
Recently, \citet{austin2014} used magnetic modelling
and rock property measurements to show that,
contrary to previous interpretations, the magnetization
of the Candelaria iron oxide copper-gold deposit,
Chile, is not dominated by the induced component.
Rather, the deposit has a relatively weak remanent
magnetization and is strongly affected by
self-demagnetization.
These examples show the importance of the self-demagnetization
and the ellipsoidal model in
producing trustworthy geological models of
high-susceptibility orebodies, which may save significant
cost associated with drilling.

A vast literature about the magnetic
modelling of ellipsoidal bodies was developed in which
are to be found the names of many researchers.
Nevertheless, interest in this subject has not yet died
out, as is evidenced by a list of modern papers in this
field.
Furthermore, the geoscientific community lacks a free easy-to-use
tool to simulate the magnetic field produced by uniformly
magnetized ellipsoids.
Such a tool could prove useful both for teaching
and researching geophysics.

In this work, we present a review of the vast literature
about the magnetic modelling of ellipsoidal bodies and a
theoretical discussion about the determination of the
isotropic susceptibility value above which
the self-demagnetization must be taken into consideration.
We propose an alternative way of determining this value
based on the body shape and the maximum relative error
allowed in the resultant magnetization.
This alternative approach is validated
by the results obtained with numerical simulations.
We also provide a set of routines to
model the magnetic field produced by ellipsoids.
The routines are written in Python language as part of
the Fatiando a Terra \citep{uieda2013}, which is an
open-source library for modelling and inversion in geophysics.
We attempt to use the best practices of continuous
integration, documentation, unit-testing and version-control
for the purpose of providing a reliable and easy-to-use
code.

@@

Let $(x, y, z)$ be a point referred to a Cartesian coordinate system with axes
$x$, $y$ and $z$ pointing to, respectively, north, east and down.
For convenience, we denominate this coordinate system as the
\textit{main coordinate system} (Fig. \ref{fig:orientation-angles}a).
Let us consider an ellipsoidal body with centre at the point
$(x_{c}, y_{c}, z_{c})$, orientation defined by the angles
\textit{strike} $\varepsilon$, \textit{dip} $\zeta$ and \textit{rake} $\eta$
(Fig. \ref{fig:orientation-angles}a), and semi-axes defined by
positive constants $a$, $b$, $c$ (Fig. \ref{fig:orientation-angles}b).
The orientation angles \textit{strike}, \textit{dip} and \textit{rake}
are commonly used to define the orientation of lines in structural geology
\citep{pollard2005, allmendinger2012}.
The points $(x, y, z)$ located on the surface of this ellipsoidal body
satisfy the following equation:
\begin{equation}
(\mathbf{r} - \mathbf{r}_c)^T \mathbf{A} (\mathbf{r} - \mathbf{r}_c) = 1 \: ,
\label{eq:ellipsoid_surface}
\end{equation}
where $\mathbf{r} = [\begin{array}{ccc} x & y & z \end{array} ]^{\top}$,
$\mathbf{r}_{c} = [\begin{array}{ccc} x_{c} & y_{c} & z_{c} \end{array} ]^{\top}$,
$\mathbf{A}$ is a positive definite matrix given by
\begin{equation}
\mathbf{A} = \mathbf{V}
\left[ \begin{array}{ccc}
a^{-2} & 0 & 0 \\
0 & b^{-2} & 0 \\
0 & 0 & c^{-2}
\end{array} \right] \mathbf{V}^{\top} \: ,
\label{eq:A}
\end{equation}
and $\mathbf{V}$ is an orthogonal matrix whose first, second and third columns
are defined by unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$ and $\mathbf{v}_{3}$
(Fig. \ref{fig:orientation-angles}b), respectively.
The matrix $\mathbf{V}$ can be defined in terms of three rotation matrices:
\begin{equation}
\mathbf{R}_{1}(\theta) = \left[ \begin{array}{ccc}
1 & 0 & 0 \\
0 & \cos \theta & \sin \theta \\
0 & -\sin \theta & \cos \theta
\end{array} \right] \: ,
\label{eq:R1}
\end{equation}
\begin{equation}
\mathbf{R}_{2}(\theta) = \left[ \begin{array}{ccc}
\cos \theta & 0 & -\sin \theta \\
0 & 1 & 0 \\
\sin \theta & 0 & \cos \theta
\end{array} \right]
\label{eq:R2}
\end{equation}
and
\begin{equation}
\mathbf{R}_{3}(\theta) = \left[ \begin{array}{ccc}
\cos \theta & \sin \theta & 0 \\
-\sin \theta & \cos \theta & 0 \\
0 & 0 & 1
\end{array} \right] \: .
\label{eq:R3}
\end{equation}
For triaxial ellipsoids (i.e. $a > b > c$) and prolate ellipsoids
(i.e. $a > b = c$), we define the orthogonal matrix $\mathbf{V}$
as follows:
\begin{equation}
\mathbf{V} =
\mathbf{R}_{1} \left( \frac{\pi}{2} \right) \;
\mathbf{R}_{2} \left( \varepsilon \right) \;
\mathbf{R}_{1} \left( \frac{\pi}{2} - \zeta \right) \;
\mathbf{R}_{3} \left( \eta \right) \: .
\label{eq:V_triaxial_prolate}
\end{equation}
For oblate ellipsoids (i.e. $a < b = c$), we define $\mathbf{V}$ as follows:
\begin{equation}
\mathbf{V} =
\mathbf{R}_{3} \left( -\frac{\pi}{2} \right) \;
\mathbf{R}_{1} \left( \pi \right) \,
\mathbf{R}_{3} \left( \varepsilon \right) \,
\mathbf{R}_{2} \left( \frac{\pi}{2} - \zeta \right) \,
\mathbf{R}_{1} \left( \eta \right) \: .
\label{eq:V_oblate}
\end{equation}
The orthogonal matrices $\mathbf{V}$ used here for triaxial, prolate and oblate
ellipsoids (Eqs. \ref{eq:V_triaxial_prolate} and \ref{eq:V_oblate}) are
different from those used by \citet{emerson1985} and \citet{clark1986}.

The magnetic modelling of an ellipsoidal body is commonly performed
in a particular Cartesian coordinate system that is aligned
with the body semi-axes
and has the origin coincident with the body centre
(Fig. \ref{fig:orientation-angles}b).
For convenience, we denominate this particular coordinate
system as the \textit{local coordinate system}.
The relationship between the Cartesian coordinates
$(\tilde{x}, \tilde{y}, \tilde{z})$ of a point in
a local coordinate system and the Cartesian
coordinates $(x, y, z)$ of the same point in the main
system is given by
\begin{equation}
\tilde{\mathbf{r}} = \mathbf{V}^{\top} \left( \mathbf{r} - \mathbf{r}_{c} \right) \: ,
\label{eq:coord_transformation}
\end{equation}
where
$\tilde{\mathbf{r}} = [\begin{array}{ccc} \tilde{x} &
                                          \tilde{y} &
                                          \tilde{z} \end{array} ]^{\top}$,
$\mathbf{r}$ and $\mathbf{r}_{c}$
are defined in Eq. \ref{eq:ellipsoid_surface} and the matrix $\mathbf{V}$
(Eq. \ref{eq:A}) is defined according to the ellipsoid type.
In what follows, quantities referred to the local coordinate system
(Fig. \ref{fig:orientation-angles}b) are indicated with
the symbol "$\sim$".

@@

Consider a magnetized ellipsoid immersed in
a uniform inducing magnetic field $\mathbf{H}_{0}$ (in $\unit{Am^{-1}}$)
given by
\begin{equation}
\mathbf{H}_{0} = \| \mathbf{H}_{0} \| \left[
\begin{array}{c}
\cos I \: \cos D \\
\cos I \: \sin D \\
\sin I
\end{array}
\right] \: ,
\label{eq:H0}
\end{equation}
where $\| \cdot \|$ denotes the Euclidean norm (or 2-norm) and
$D$ and $I$ are respectively the declination and inclination of the
local geomagnetic field in the main coordinate system
(Fig. \ref{fig:orientation-angles}a).
This field represents the main component of the
Earth's magnetic field, which is usually assumed to be generated
by the Earth's liquid core.
In the absence of conduction currents,
the total magnetic field $\mathbf{H}(\mathbf{r})$
at the position $\mathbf{r}$ (Eq. \ref{eq:ellipsoid_surface})
of a point referred to the main
coordinate system is defined as follows \citep{sharma1966,
eskola1980, reitz1992, sttraton2007}:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0} - \nabla V(\mathbf{r}) \: ,
\label{eq:H}
\end{equation}
where the second term is the negative gradient of
the magnetic scalar potential $V(\mathbf{r})$ given by
\begin{equation}
V(\mathbf{r}) = -\frac{1}{4\pi} \iiint_{\vartheta}
\mathbf{M}(\mathbf{r}^{\prime})^{\top}
\nabla \left(
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\right) \, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:V-potential}
\end{equation}
In this equation, $\mathbf{r}^{\prime} = [\begin{array}{ccc}
x^{\prime} & y^{\prime} & z^{\prime} \end{array} ]^{\top}$
is the position vector of a point located within the volume $\vartheta$,
the integral is conducted over the variables
$x^{\prime}$, $y^{\prime}$ and, $z^{\prime}$ and
$\mathbf{M}(\mathbf{r}^{\prime})$ is the magnetization vector
(in \unit{Am^{-1}}).
Eq. \ref{eq:V-potential} is valid anywhere,
independently if the position vector $\mathbf{r}$ represents
a point located inside or outside the magnetized body
\citep{dubois1896, reitz1992, sttraton2007}.

Based on \citeauthor{maxwell1873}'s postulate, let us
assume that the body has a uniform magnetization given by
\begin{equation}
\mathbf{M} = \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:M-KH}
\end{equation}
where $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field
at any point within the body and $\mathbf{K}$ is a
constant and symmetrical second-order tensor representing the
magnetic susceptibility of the body.
This is a good approximation for bodies at room temperature,
subjected to an inducing field $\mathbf{H}_{0}$ with
strength $\leq 10^{-3} \mu_{0}^{-1}$ \unit{A \, m^{-1}} \citep{rochette1992},
where $\mu_{0}$ represents the magnetic constant (in \unit{H \, m^{-1}}).
In this case, the susceptibility tensor $\mathbf{K}$ is commonly
represented, in the main coordinate system (Fig. \ref{fig:orientation-angles}a),
as follows:
\begin{equation}
\mathbf{K} = \mathbf{U}
\left[ \begin{array}{ccc}
k_{1} & 0 & 0 \\
0 & k_{2} & 0 \\
0 & 0 & k_{3}
\end{array} \right] \mathbf{U}^{\top} \: ,
\label{eq:K}
\end{equation}
where $k_{1} > k_{2} > k_{3}$ are the
\textit{principal susceptibilities} and $\mathbf{U}$ is
an orthogonal matrix whose columns $\mathbf{u}_{i}$,
$i = 1, 2, 3$, are unit vectors called
\textit{principal directions}.
Similarly to the matrix $\mathbf{V}$ (Eqs. \ref{eq:ellipsoid_surface},
\ref{eq:V_triaxial_prolate} and \ref{eq:V_oblate}), we define
the matrix $\mathbf{U}$ as a function
of given orientation angles $\varepsilon$, $\zeta$
and $\eta$ depending on the ellipsoid type. For triaxial
and prolate ellipsoids, we define $\mathbf{U}$ by
using Eq. \ref{eq:V_triaxial_prolate}, whereas for oblate ellipsoids
we use Eq. \ref{eq:V_oblate}.
Notice that the orientation angles $\varepsilon$, $\zeta$,
and $\eta$ defining the orthogonal matrix $\mathbf{U}$
may be different from the angles $\varepsilon$, $\zeta$
and $\eta$ defining the ellipsoid orientation
(Fig. \ref{fig:orientation-angles}).

If the principal susceptibilities are different from
each other, we say that the body has an
anisotropy of magnetic susceptibility (AMS).
The AMS is generally associated with the preferred orientation
of the grains of magnetic minerals forming the rock
\citep{fuller1963, uyeda1963, janak1972, hrouda1982,
thompson1986, macdonald1987, rochette1992, dunlop1997,
tauxe2003}.
For the particular case in which the principal directions
coincide with the ellipsoid axes, the matrix $\mathbf{U}$ is
equal to the matrix $\mathbf{V}$ (Eq. \ref{eq:A}).
Another important particular case is that in which the
susceptibility is isotropic and, consequently, the principal
susceptibilities $k_{1}$, $k_{2}$, and $k_{3}$ (Eq. \ref{eq:K})
are equal to a constant $\chi$. In this case, the susceptibility
tensor $\mathbf{K}$ (Eq. \ref{eq:K}) assumes the particular form
\begin{equation}
\mathbf{K} = \chi \, \mathbf{I} \: ,
\label{eq:K-isotropic}
\end{equation}
where $\mathbf{I}$ represents the identity matrix.
